Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-21 appendix-EPQ.tex
modify

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
  \label{fig:automata for FIFO-1 in appendix}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.

@@

Let $\textit{Auts}_{\textit{sinPri}} = \{ \mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4 \}$. Let us prove Lemma \ref{lemma:automata for extended priority queue with single priority}.

\begin{lemma}\label{lemma:automata for extended priority queue with single priority}
Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
\end{lemma}

@@

\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and at least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.

Let us prove the $\textit{only if}$ direction. Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by an automaton in $\textit{Auts}_{\textit{sinPri}}$. By data-independence, we can see that there exists a data-differentiated $e \in \mathcal{I}$ and renaming function, such that $e_0=r(e)$. Let $e'$ be obtained from $e$ by first removing $\textit{rm}(\textit{empty})$, and then,

@@

Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by some automaton in $\textit{Auts}_{\textit{sinPri}}$. This completes the proof of this lemma. \qed
\end {proof}

@@

Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a witness automaton.

For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}. Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $c_4 = c_1 + \textit{ret}(\textit{put},b)$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

\begin{restatable}{lemma}{MaximalInPBadGPMakePQ1Equal}
\label{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$. If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$. Then $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
\end{restatable}

@@

\item[-] For item $z$ with priority smaller than $\textit{pri}$. If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions. If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions. Otherwise, $x$ is in interval of $z$, which contradicts the definition of gap-point, and is impossible.
\end{itemize}

Let $l$ be the sequence of linearization points constructed above. It is obvious that $e \sqsubseteq l$. Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$. It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$. Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$. Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$. Since $\textit{last}(e) = \seqPQ_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.

@@

According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},x)$.

@@

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$. We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$. Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.

@@

    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.

    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$. If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$. Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.

\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$. We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
    By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.

@@

    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$. Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$. Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
    \end{itemize}
\end{itemize}

@@

    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

@@

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^C c_1$. Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.

    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$. Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. It is obvious that $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

@@

\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
\end{itemize}

>>>>

We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
  \label{fig:automata for FIFO-1 in appendix}
\end{figure}

We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\top) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,\top), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.

@@

Let $\mathcal{A}_{\textit{sinPri}}$ be the union of $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4$. Let us prove Lemma \ref{lemma:automata for extended priority queue with single priority}.

\begin{lemma}\label{lemma:automata for extended priority queue with single priority}
Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \mathcal{A}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
\end{lemma}

@@

\cite{DBLP:conf/icalp/BouajjaniEEH15} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and at least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.

Let us prove the $\textit{only if}$ direction. Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{SinPri}}^4$. By data-independence, we can see that there exists a data-differentiated $e \in \mathcal{I}$ and renaming function, such that $e_0=r(e)$. Let $e'$ be obtained from $e$ by first removing $\textit{rm}(\textit{empty})$, and then,

@@

Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{SinPri}}^4$. This completes the proof of this lemma. \qed
\end {proof}

@@

Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a register automaton.

For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}. Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $c_4 = c_1 + \textit{ret}(\textit{put},b)$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

\begin{restatable}{lemma}{MaximalInPBadGPMakePQ1Equal}
\label{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$. If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},y)$. Then $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
\end{restatable}

@@

\item[-] For item $z$ with priority smaller than $\textit{pri}$. If both $\textit{call}(\textit{put},z,\_)$ and $\textit{call}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions. If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions. Otherwise, $x$ is in interval of $z$, which contradicts the definition of gap-point, and is impossible.
\end{itemize}

Let $l$ be the sequence of linearization points constructed above. It is obvious that $e \sqsubseteq l$. Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$. It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$. Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$. Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$. Since $\textit{last}(e) = \seqPQ_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.

@@

According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.

@@

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{call}(\textit{put},y,\textit{pri})$ and $\textit{call}(\textit{rm},x)$. We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$. Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.

@@

    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$.

    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$. If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$. Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.

\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{call}(\textit{put},a_2,\textit{pri})$ or $\textit{call}(\textit{rm},a_2)$ in $e$. We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
    By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.

@@

    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,\textit{pri})$ and $\textit{call}(\textit{rm},a_2)$. Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$. Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
    \end{itemize}
\end{itemize}

@@

    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

@@

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^C c_1$. Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.

    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$. Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$. It is obvious that $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

@@

\item[-] $\textit{act}_2$ is before $\textit{call}(\textit{rm},d_1)$.
\end{itemize}
